Official Development Assistance and Institutional Quality on Undeveloped countries

### 2024-08-24 paper.Rmd
comentarios y html actualizado

---
title: "Official Development Assistance and Institutional Quality on Undeveloped countries"
output:
  pdf_document: default
  html_notebook: default
date: "2024-08-05"
author: "Oscar Eduardo Morales Cárdenas"
---

@@

\begin{flushright}
Watts, Carl. (2014). Re: Does foreign aid help the developing countries towards development?. Retrieved from: $https://www.researchgate.net/post/Does_foreign_aid_help_the_developing_countries_towards_development/5322005ed039b1e7648b459c/citation/download.$
\end{flushright}

@@

\begin{flushright}
Ekanayake, E. & Chatrna, Dasha. (2010). The effect of foreign aid on economic growth in developing countries. Journal of International Business and Cultural Studies. 3.
\end{flushright}

@@

\begin{flushright}
Hayaloğlu, Pınar. (2023). Foreign Aid, Institutions, and Economic Performance in Developing Countries. Eskişehir Osmangazi Üniversitesi İktisadi ve İdari Bilimler Dergisi. 18. 748-765. 10.17153/oguiibf.1277348. 
\end{flushright}

# Loading libraries

Some libraries and packages used for data manipulation and data scrapping

```{r}
library(tidyverse) # i don't think I´ll use this /r
library(WDI)       # for World Bank data acceding (mostly country code names)
library(readxl)    # for excel files reading
library(readr)     # for csv files reading
library(visdat)    # for data visualization
library(plotly)    # for plots
library(purrr)     # for map funciton
```

# Gathering Data

Data for low income countries will be used, as categorized by the World Bank there are 26 Low income countries and 51 Lower middle income countries

@@

Here are listed countries to use:

```{r}
my_countries <- country_class %>%
  filter(!is.na(Region), `Income group` %in% c('Low income', 'Lower middle income')) %>%
  select(Code)
```

Here we get the respective iso2c names

```{r}
my_countries$iso2c <- WDI_data$country %>%
  filter(iso3c %in% my_countries$Code) %>%
  .$iso2c
```

Data from the World Bank API and the Human Development Reports API is dowloaded by the usage of Python Scripts.
They are stored as csv files and then loaded here:

@@

# Data Manipulation

Transform the data into a new structure for easier understanding

@@

Lets check which data is missing...

```{r}
vis_dat(datos_paper %>% select(all_of(gsub("_", ".", oda_indicators)))) 
  # DT.ODA.OATL.CD and DT.ODA.OATL.KD missing for all countries and years
  # DT.ODA.ODAT.GI.ZS, DT.ODA.ODAT.GN.ZS, DT.ODA.ODAT.MP.ZS and DT.ODA.ODAT.XP.ZS also has some missing values
  # There is a couple of countries that has some missing values for some exact years
```
```{r}
vis_dat(datos_paper %>% arrange(year) %>%
          select(all_of(gsub("_", ".", gob_indicators)))) 
  # It seems that some years are missing here, should we take from (almost) the half and upward? 
```
```{r}
vis_dat(datos_paper %>%
          select(all_of(hdi_indicators))) 
  # abr, co2_prod, le, le_f, le_m, mmr seems like has all the years, but... doesn't hdi? 
```

```{r}
vis_dat(datos_paper %>% arrange(year) %>% select(hdi)) 
  # hdi missing in various countries and years... what do we do?
```
Taking into account the missing data let´s select a smaller sample to test
```{r}
datos_paper %>% filter(is.na(DT.ODA.ALLD.CD)) ## Culprit is SS (South Sudan) and ZW (Zimbabwe) they have almost no ODA and
                                              ## GOB indicators
datos_paper %>% filter(!iso2c %in% c('SS', 'ZW')) %>% filter(is.na(CC.EST)) %>% group_by(year) %>% summarise(times = n())
  # it seems that 1995, 1997, 1999 and 2001 didn't measure governance indicators at all
  # 1996, 1998, 2000, 2002 and 2003 has some missing countries, it seems lets take a look
datos_paper %>% arrange(year) %>% filter(!iso2c %in% c('SS', 'ZW'), !year %in% c(1995, 1997, 1999, 2001)) %>%
                filter(is.na(CC.EST)) # FM (Micronesia), KI (Kiribati) and TL (Timor-Leste) didnt have GOB in those years
                                      # also CV (Cabo Verde) and SB (Solomon Islands) didn't register some GOB in 2000 - 03
```
Now lets view data again with those filters
```{r}
datos_paper %>% 
          arrange(iso2c) %>% 
          filter(!iso2c %in% c('SS', 'ZW', 'BT', 'ER', 'GW', 'KP', 'LB', 'NG', 'PS', 'SO', 'VU', 'FM', 'KI', 'TL'), 
                 !year %in% c(1995, 1996, 1997, 1998, 1999, 2001)) %>%
          select(iso2c, year, hdi,
                 all_of(gsub("_", ".", gob_indicators))
                 ) %>%
         filter(is.na(GE.EST)) #%>%
         # group_by(year) %>%
         # summarise(times = n()) 
# 1996 and 1998 are troublesome, so lets start in 2000
# BT (Bhutan), ER (Eritrea), GW (Guinea-Bissau), KP (North Korea), LB (Lebanon), NG (Nigeria), PS (Palestine), SO (Somalia),
# VU (Vanuatu) are the countries without hdi
```

```{r}
vis_dat(datos_paper %>% 
        filter(!iso2c %in% c('SS', 'ZW', 'BT', 'ER', 'GW', 'KP', 'LB', 'NG', 'PS', 'SO', 'VU', 'FM', 'KI', 'TL', 'CV', 'SB'),
               !year %in% c(1995, 1996, 1997, 1998, 1999, 2000, 2001)) %>%
        select(iso2c, year, hdi, DT.ODA.ALLD.CD, DT.ODA.ALLD.KD, DT.ODA.ODAT.CD, DT.ODA.ODAT.KD, DT.ODA.ODAT.PC.ZS,
               all_of(gsub("_", ".", gob_indicators))
               ))
```
From 1925 to 1098 rows huh... i don't know...
```{r}
datos_model <- datos_paper %>% 
        filter(!iso2c %in% c('SS', 'ZW', 'BT', 'ER', 'GW', 'KP', 'LB', 'NG', 'PS', 'SO', 'VU', 'FM', 'KI', 'TL', 'CV', 'SB'),
               !year %in% c(1995, 1996, 1997, 1998, 1999, 2000, 2001)) %>%
        select(iso2c, year, hdi, DT.ODA.ALLD.CD, DT.ODA.ALLD.KD, DT.ODA.ODAT.CD, DT.ODA.ODAT.KD, DT.ODA.ODAT.PC.ZS,
               all_of(gsub("_", ".", gob_indicators))
               )
```

Let's give it a shot with a linear model
```{r}
model <- lm(hdi ~ DT.ODA.ALLD.CD + CC.EST + GE.EST + PV.EST + RQ.EST + RL.EST, data=datos_model)
summary(model)
```
It worked!!!
but I just threw random variables in there, I want to see their relationship with HID

#Data visualization

@@

... what was that???
I guess I cant just throw things in there and expect to have a good model, I can't see a clear relationship between this stuff, more than better governance indicators are usually accompanied with better HDI scores.
But... I was saying that I wanted to find a relationship with HDI changes, not the actual HDI score. So, I'll compute a HDI change per year per country, (and lose 61 rows in the process, yay...)

# Data manipulation 2 Electric boogaloo

Note: originally I was excluding data from 2001, because some countries didn't have governance indicators for that year particularly, but!, since I'll lose one year by computing the growth in hdi I can still use this year and wont loose data al all. STONKS!!!

@@

vis_dat(datos_model_3)
```
and here we go again
```{r}
my_plot_3  <- list()

@@

```{r}
subplot(my_plot_3[6:11], nrows = 2, margin = 0.05) %>% layout(title = 'CC vs HDI diff')
```
```{r}
subplot(my_plot_3[12:17], nrows = 2, margin = 0.05) %>% layout(title = 'GE vs HDI diff')
```

@@

Testing model while using hdi_diff as dependent variable (it went better than expected)
```{r}
model_3 <- lm(hdi_diff ~ DT.ODA.ALLD.CD + CC.EST + GE.EST + PV.EST + RQ.EST + RL.EST, data=datos_model_3)
summary(model_3)
```
I liked the relationship visible in the PER.RNK gov indicators, let's try a model with those variables (it went worse than expected)
```{r}
model_4 <- lm(hdi_diff ~ DT.ODA.ALLD.CD + CC.PER.RNK + GE.PER.RNK + PV.PER.RNK + RQ.PER.RNK + RL.EST, data=datos_model_3)
summary(model_4)
```

>>>>

---
title: "Official Development Assistance and Institutional Quality on Undeveloped countries"
author: "Oscar Eduardo Morales Cárdenas"
date: "2024-08-05"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

@@

\textbf{Watts, Carl. (2014). Re: Does foreign aid help the developing countries towards development?. Retrieved from: $https://www.researchgate.net/post/Does_foreign_aid_help_the_developing_countries_towards_development/5322005ed039b1e7648b459c/citation/download.$}

@@

\textbf{Ekanayake, E. & Chatrna, Dasha. (2010). The effect of foreign aid on economic growth in developing countries. Journal of International Business and Cultural Studies. 3.}

@@

\textbf{Hayaloğlu, Pınar. (2023). Foreign Aid, Institutions, and Economic Performance in Developing Countries. Eskişehir Osmangazi Üniversitesi İktisadi ve İdari Bilimler Dergisi. 18. 748-765. 10.17153/oguiibf.1277348.}

# Cargando Librerias

Algunas librerias y paquetes usados para obtener y descargar los datos

```{r}
library(tidyverse) # manejo de dataframes
library(WDI)       # libreria para acceder a metadata de banco mundial
library(readxl)    # leer archivos de excel
library(readr)     # leer archivos csv
library(visdat)    # visualizacion de datos como graficos
library(plotly)    # graficos
library(purrr)     # funcion map
```

# Obtener datos

Datos para paises bajos ingresos sean utilizados, segun clasificación del banco mundial, hay 26 paises de bajos ingresos y 51 de ingresos medios bajos

@@

Listado de paises a analisar:

```{r}
my_countries <- country_class %>%
  filter(!is.na(Region), `Income group` %in% c('Low income', 'Lower middle income')) %>%
  select(Code)
my_countries
```

Hacer la respectiva asociacion de nombres iso3c e iso2c

```{r}
my_countries$iso2c <- WDI_data$country %>%
  filter(iso3c %in% my_countries$Code) %>%
  .$iso2c
my_countries
```

Datos del banco mundial (para ODA y los indices de gobernanza) y el Human Development Reports API son descargados desde scripts de Python.
Son almacenados en archivos CSV y luego son cargados aqui:

@@

# Manipulacion de Datos

Transformar la estructura de los datos para una mejor comprension

@@

Revisar que datos estan como faltantes

```{r}
vis_dat(datos_paper %>% select(all_of(gsub("_", ".", oda_indicators)))) 
  # DT.ODA.OATL.CD and DT.ODA.OATL.KD faltan
  # DT.ODA.ODAT.GI.ZS, DT.ODA.ODAT.GN.ZS, DT.ODA.ODAT.MP.ZS and DT.ODA.ODAT.XP.ZS tienen faltas
  # Un par de ocurrencias pais-año que faltan datos
```

```{r}
vis_dat(datos_paper %>% arrange(year) %>%
          select(all_of(gsub("_", ".", gob_indicators)))) 
  # Datos del 2000 para atras tienen espacios faltantes 
```

```{r}
vis_dat(datos_paper %>%
          select(all_of(hdi_indicators))) 
  # abr, co2_prod, le, le_f, le_m, mmr son las pocas categorias sin datos faltantes
```

```{r}
vis_dat(datos_paper %>% arrange(year) %>% select(hdi)) 
  # hdi faltante en multiples ocaciones
```

Tomando en cuenta los datos faltantes, hacer filtros para seleccionar una muestra mas pequeña
```{r}
datos_paper %>% filter(is.na(DT.ODA.ALLD.CD)) ## SS (South Sudan) y ZW (Zimbabwe) faltantes de ODA y GOB indicators

datos_paper %>% filter(!iso2c %in% c('SS', 'ZW')) %>% filter(is.na(CC.EST)) %>% group_by(year) %>% summarise(times = n())
  # para años 1995, 1997, 1999 y 2001 no hay registros de GOB
  # 1996, 1998, 2000, 2002 and 2003 tiene algunos paises sin datos
datos_paper %>% arrange(year) %>% filter(!iso2c %in% c('SS', 'ZW'), !year %in% c(1995, 1997, 1999, 2001)) %>%
                filter(is.na(CC.EST)) # FM (Micronesia), KI (Kiribati) y TL (Timor-Leste) no tiene GOB in en estos años 
                                      # tambien CV (Cabo Verde) and SB (Solomon Islands) no registro GOB en 2000 - 2003
```
Ver datos aplicando los filtros determinados en las busquedas pasadas
antes del 2001 suele tener informacion faltante
BT (Bhutan), ER (Eritrea), GW (Guinea-Bissau), KP (North Korea), LB (Lebanon), NG (Nigeria), PS (Palestine), SO (Somalia), VU (Vanuatu) son paises sin registro de hdi
```{r}
datos_paper %>% 
          arrange(iso2c) %>% 
          filter(!iso2c %in% c('SS','ZW','BT','ER','GW','KP','LB','NG','PS','SO','VU','FM','KI','TL','CB','CV','SB'), 
                 !year %in% c(1995, 1996, 1997, 1998, 1999, 2000, 2001)) %>%
          select(iso2c, year, hdi,
                 all_of(gsub("_", ".", gob_indicators))
                 )

```{r}
vis_dat(datos_paper %>% 
        filter(!iso2c %in% c('SS', 'ZW', 'BT', 'ER', 'GW', 'KP', 'LB', 'NG', 'PS', 'SO', 'VU', 'FM', 'KI', 'TL', 'CV', 'SB'),
               !year %in% c(1995, 1996, 1997, 1998, 1999, 2000, 2001)) %>%
        select(iso2c, year, hdi, DT.ODA.ALLD.CD, DT.ODA.ALLD.KD, DT.ODA.ODAT.CD, DT.ODA.ODAT.KD, DT.ODA.ODAT.PC.ZS,
               all_of(gsub("_", ".", gob_indicators))
               ))
```
De 1925 observaciones reducimos a 1098
```{r}
datos_model <- datos_paper %>% 
        filter(!iso2c %in% c('SS', 'ZW', 'BT', 'ER', 'GW', 'KP', 'LB', 'NG', 'PS', 'SO', 'VU', 'FM', 'KI', 'TL', 'CV', 'SB'),
               !year %in% c(1995, 1996, 1997, 1998, 1999, 2000, 2001)) %>%
        select(iso2c, year, hdi, DT.ODA.ALLD.CD, DT.ODA.ALLD.KD, DT.ODA.ODAT.CD, DT.ODA.ODAT.KD, DT.ODA.ODAT.PC.ZS,
               all_of(gsub("_", ".", gob_indicators))
               )
```

Probando un modelo sencillo Minimos cuadrados
```{r}
model <- lm(hdi ~ DT.ODA.ALLD.CD + CC.EST + GE.EST + PV.EST + RQ.EST + RL.EST, data=datos_model)
summary(model)
```
Todas las variables son significativas al 95% excepto ODA

Se revisara las relaciones entre las variables graficamente

#Visualizacion de datos

@@

No se ve una relacion clara, hay tanto paises con punteos altos y bajos de GOB que tienen tanto HID altos o bajos
Quiza puede verse una leve relacion de mayor punteo en GOB acompañado de mejor punteo den HDI

Se realizara el mismo proceso con el crecimiento o decrecimiento de HDI anual (no se perderan datos al calcular la diferencia porque se añade el año 2001 en la seleccion)

# Manipulacion de datos utilizando operador diferencia

@@

vis_dat(datos_model_3)
```
Construccion de graficos
```{r}
my_plot_3  <- list()

@@

```{r}
subplot(my_plot_3[6:11], nrows = 2, margin = 0.05) %>% layout(title = 'CC vs HDI diff')
```

```{r}
subplot(my_plot_3[12:17], nrows = 2, margin = 0.05) %>% layout(title = 'GE vs HDI diff')
```

@@

Probar un nuevo modelo usadno ahora como variable dependiente hdi_diff
```{r}
model_3 <- lm(hdi_diff ~ DT.ODA.ALLD.CD + CC.EST + GE.EST + PV.EST + RQ.EST + RL.EST, data=datos_model_3)
summary(model_3)
```
Otro modelo ahora usando las variables PER.RNK en vex de EST
```{r}
model_4 <- lm(hdi_diff ~ DT.ODA.ALLD.CD + CC.PER.RNK + GE.PER.RNK + PV.PER.RNK + RQ.PER.RNK + RL.EST, data=datos_model_3)
summary(model_4)
```
Viendo la historia de las variables en el tiempo (por pais)
```{r}
datos_model_3 %>% filter(iso2c == 'AF') %>% plot_ly(x = ~year) %>% 
  add_trace(y = ~hdi, type = 'scatter', mode = 'lines+markers', name = 'hdi') %>% 
  add_trace(y = ~hdi_diff, type = 'scatter', mode = 'lines+markers', name = 'hdi diff') %>% 
  add_trace(y = ~DT.ODA.ALLD.CD / 10000000000, type = 'scatter', mode = 'lines+markers', name = 'ODA')  %>% 
  add_trace(y = ~CC.EST, type = 'scatter', mode = 'lines+markers', name = 'CC') %>% 
  add_trace(y = ~GE.EST, type = 'scatter', mode = 'lines+markers', name = 'GE')  %>% 
  add_trace(y = ~PV.EST, type = 'scatter', mode = 'lines+markers', name = 'PV')  %>% 
  add_trace(y = ~RQ.EST, type = 'scatter', mode = 'lines+markers', name = 'RQ')  %>% 
  add_trace(y = ~RL.EST, type = 'scatter', mode = 'lines+markers', name = 'RL') 
```
